<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{GENE.E Overview}
-->

```{r setup, echo=FALSE, warn=FALSE}
library(knitr)
options(width=80)
```
```{r wrap-hook, echo=FALSE}
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

# GENE.E

The `GENE.E` package provides the ability to transfer matrices, dendrograms, and annotations between R and [GENE-E](www.broadinstitute.org/cancer/software/GENE-E/). Two methods are provided:
*	from.genee Gets the current selection from GENE-E.
*	to.genee Sends a matrix and annotations to GENE-E.

# Example Usage

For illustration of `GENE.E` usage, we use the Golub et al. data
set on acute leukemia from the `golubEsets` package. First, load the `GENE.E` and `golubEsets` packages:

```{r library_GENE.E_fake, eval=FALSE}
library(GENE.E)
library(golubEsets) 
```
```{r library_GENE.E_real, warn=FALSE, echo=FALSE, results="hide"}
suppressPackageStartupMessages(library(GENE.E))
suppressPackageStartupMessages(library(golubEsets))
```
Next, ensure that GENE-E is running and then issue the following commands to load the data and send the data to GENE-E:

```{r}
data(Golub_Merge) 
to.genee(exprs(Golub_Merge), fData(Golub_Merge), show.colnames=F, pData(Golub_Merge)[,c('Gender', 'ALL.AML')], column.hclust=hclust(dist(t(exprs(Golub_Merge)))))
```

In GENE-E, (see video below), drag the dotted line to cut the sample dendrogram. Next, click a dendrogram branch to select the samples in that branch. 

<iframe width="853" height="480" src="http://www.youtube.com/embed/z_CUlLlkmGI?rel=0" frameborder="0" allowfullscreen></iframe>

Finally, get the selected data in GENE-E back into our R session:

```{r}
Sys.sleep(5)
selection <- from.genee()
```